---
title: "BIS-15L-Project"
author: "Eva Moncada"
date: "2024-03-02"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Installing the Necessary Packages
```{r}
#install.packages("RColorBrewer")
#install.packages("paletteer")
#install.packages("ggthemes")
#install.packages("ggmap")
```

# Loading the Necessary Libraries
```{r}
library("ggplot2")
library("naniar")
library("tidyverse")
library("janitor")
library("RColorBrewer")
library("paletteer")
library("lubridate")
library("ggmap")
```

# Loading the crime dataset
### Note: The file was too bog got GitHub so I had to delete it from my data folder under the repository. Here is the [link](https://catalog.data.gov/dataset/crime-data-from-2020-to-present) to the website where the data came from for your reference:

```{r}
original_crime_data <- read_csv("data/Crime_Data_from_2020_to_Present.csv") %>%
  clean_names()
```

##Cleaning the NAs
The NA's in this dataset are represented by blank cells.

```{r}
cleaned_crime <- original_crime_data %>%
  mutate_all(~ifelse(. == "", NA, .))
crime<-cleaned_crime
crime
```

##Taking a look at the structure, dimensions, and name of the crime data
```{r}
glimpse(cleaned_crime)
```

```{r}
dim(crime)
```

```{r}
names(crime)
```


# Which Age Group Commits the Most Crimes In Los Angeles County?
Presented on Slide 11 of the presentation, we will be observing which age group is shown to have committed the most crimes. Furthermore, we are interested in observing which crimes were committed the most in the county. Lastly, the year with the most crimes may provide insight as to why the crimes may have been commited, however these are all hypothetical (ex. if it was during COVID years, it could be due to financial struggles)

## Determining the Most Common Age Group
```{r}
crime %>%
  group_by(vict_age) %>%
  summarise(crime_count = n()) %>%
  arrange(desc(crime_count))
```

vict_age = 0 most likely represents data that was not found and would be considered an NA. 

## Plotting the Age Groups

### Plotting by Individual Ages - Did Not Use in Presentation 
```{r, fig.width=10, fig.height=6}
crime %>%
  filter(vict_age > 14 & vict_age < 100) %>%
  ggplot(aes(x = vict_age)) +
  geom_histogram(binwidth = 1, fill = "#FFA07A", color = "black") + 
  labs(title = "Distribution of Crime Counts by Age",
       x = "Age",
       y = "Frequency") +
  scale_x_continuous(breaks = seq(15, 100, by = 5)) +
  theme_bw() +
  theme(text = element_text(family = "Courier"),
        panel.background = element_rect(fill = "#FDF7F6"))


```

This last graph looked a bit too intense. So, out of curiosity, I wanted to see how each age interval compare to each other in the number of crime incidents committed. 

### Plotting Age Intervals - Slide 9
```{r}
crime %>%
  filter(vict_age > 14) %>%
  mutate(age_group = cut(vict_age, 
                        breaks = c(14, 20, 30, 40, 50, 60, 70, 80, 90, Inf), 
                        labels = c("14-19", "20-29", "30-39", "40-49", "50-59", "60-69", "70-79", "80-89", "90+"),
                        include.lowest = TRUE)) %>%
  group_by(age_group) %>%
  summarise(crime_count = n()) %>%
  arrange(desc(crime_count))
```

## Creating the plot 
```{r}
crime %>%
  filter(vict_age > 14) %>%
  mutate(age_group = cut(vict_age, 
                        breaks = c(14, 20, 30, 40, 50, 60, 70, 80, 90, Inf), 
                        labels = c("14-19", "20-29", "30-39", "40-49", "50-59", "60-69", "70-79", "80-89", "90+"),
                        include.lowest = TRUE)) %>%
  group_by(age_group) %>%
  summarise(crime_count = n()) %>%
  arrange(desc(crime_count)) %>%
  ggplot(aes(x = age_group, y = crime_count, fill = age_group)) +
  geom_bar(stat = "identity", color = "black") +
  geom_text(aes(label = crime_count), vjust = -0.5, color = "black", size = 3, family = "Courier") +
  labs(title = "Crime Counts by Age Group",
       x = "Age Group",
       y = "Crime Count") +
  scale_fill_manual(values = c("#FFDAB9", "#D2B48C", "#FFA07A", "#FFE4C4", "#FFB6C1", "#FFD700", "#FAF0E6", "#FF6347", "#F5DEB3")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6), 
        panel.background = element_rect(fill = "#FDF7F6"),
        text = element_text(family = "Courier"),  
        legend.text = element_text(family = "Courier"),  
        legend.title = element_text(family = "Courier Bold")) +
  labs(fill = "Age Group")
```

As a group, it seemed like 20-29 year old as a whole committed the most number of crimes, though 30 year olds had the most when comparing individuals ages. 

# Observing which Crimes were Committed the Most

## What year was most of these crimes committed? Is there a significant increase in one of the years? Why do you think that is?

### Creating a separate column in the dataset for "year" so I can observe the changes per year. 
```{r}
crime <- crime %>%
  mutate(year_occurred = year(mdy_hms(date_occ)))
```

```{r}
crime %>%
  group_by(year_occurred) %>%
  summarise(crime_count = n())
```

## Plotting the Total Crime Incidents Over the Years - Slide 10
```{r}
crime %>%
  group_by(year_occurred) %>%
  summarise(crime_count = n()) %>%
  ggplot(aes(x = as.factor(year_occurred), y = crime_count, fill = as.factor(year_occurred))) +
  geom_bar(stat = "identity", color = "black") +
  geom_text(aes(label = crime_count), vjust = -0.5, color = "black", size = 3, family = "Courier") + 
  labs(title = "Total Crime Incidents per Year",
       x = "Year",
       y = "Crime Count",
       fill = "Year") +
  scale_fill_manual(values = c("#FFDAB9", "#D2B48C", "#FFA07A", "#FFE4C4", "#FFB6C1", "#FFD700", "#FAF0E6", "#FF6347", "#F5DEB3")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6, family = "Courier"),
        panel.background = element_rect(fill = "#FDF7F6"),
        text = element_text(family = "Courier"),
        legend.text = element_text(family = "Courier"),  
        legend.title = element_text(family = "Courier Bold"))

```

# Most Common Crimes - (Netra Ended up Presenting This so I cut it from my slides)
```{r, fig.width=15, fig.height=8}
crime %>%
  group_by(crm_cd_desc) %>%
  summarise(crime_count = n(), .groups = 'drop') %>%
  arrange(desc(crime_count)) %>%
  slice_head(n = 10) %>%
  na.omit() %>% 
  ggplot(aes(x = reorder(crm_cd_desc, crime_count), y = crime_count, fill = crm_cd_desc)) +
  geom_bar(stat = "identity", color = "black") +
  geom_text(aes(label = crime_count), vjust = -0.5, color = "black", size = 3, family = "Courier") +
  labs(title = "Top 10 Crimes Committed between 2020-2024",
       x = "Crime Description",
       y = "Crime Count",
       fill = "Crime Type") +
  scale_fill_manual(values = c("#FFDAB9", "#D2B48C", "#FFA07A", "#FFE4C4", "#FFB6C1", "#FAF0E6", "#FF6347", "#F5DEB3", "#CD853F", "#E9967A", "#F4A460", "#E9967A", "#FF4500")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6, family = "Courier"),
        panel.background = element_rect(fill = "#FDF7F6"),
        text = element_text(family = "Courier"),
        legend.text = element_text(family = "Courier"),
        legend.title = element_text(family = "Courier Bold"))

```

# Weaponry

One of the more interesting parts of observing crimes is analyzing the weapon of choice and what that weapon was used for. When determining the weapon of choice, connecting it to the most common crime type will allow us to deduce why that weapons was chosen. (ex. in murders, are knives more common than guns? If yes, then why? Is it because it is easier to find in the common household/easier to get a hold of?)

## How Often Were Weapons Used?
```{r}
crime %>%
  mutate(weapon_status = ifelse(!is.na(weapon_used_cd) & weapon_used_cd != "None", "Weapon Used", "Weapon Not Used")) %>%
  group_by(weapon_status) %>%
  summarise(crime_count = n())
```

##Plotting the Pie Chart - Slide 11
```{r}
color_palette <- c("Weapon Used" = "#D2B48C", "Weapon Not Used" = "#FFA07A", "Unsure" = "#FDF7F6")
crime_summary <- crime %>%
  mutate(weapon_status = ifelse(!is.na(weapon_used_cd) & weapon_used_cd != "None", "Weapon Used", "Weapon Not Used")) %>%
  group_by(weapon_status) %>%
  summarise(crime_count = n(), .groups = 'drop') %>%
  mutate(percentage = crime_count / sum(crime_count) * 100)
ggplot(crime_summary, aes(x = "", y = crime_count, fill = weapon_status, label = paste0(round(percentage), "%"))) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar("y", start = 0) +
  geom_text(position = position_stack(vjust = 0.5), size = 3, family = "Courier") +
  labs(title = "Crimes by Use of Weaponry",
       fill = "Weapon Status") +
  scale_fill_manual(values = color_palette) +
  theme_void() +
  theme(legend.position = "right",
        text = element_text(family = "Courier"))
```
We can see that 65% of crimes did not use weapons. Conversely, 35% of crimes had a weapon involved. We want to observe which weapons were most commonly used in Los Angeles.

## Determining the Most Used Weapon
```{r}
crime %>%
  group_by(weapon_desc) %>%
  summarise(crime_count = n()) %>%
  arrange(desc(crime_count))
```

## Plotting the Top 10 Most Used Weaponry - Slide 12
```{r}
crime %>%
  filter(!is.na(weapon_desc)) %>%
  count(weapon_desc) %>%
  arrange(desc(n)) %>%
  slice_head(n = 10) %>%
  mutate(weapon_desc = if_else(row_number() == 11, "Other", as.character(weapon_desc))) %>%
  ggplot(aes(x = "", y = n, fill = weapon_desc)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar(theta = "y") +
  labs(title = "Top Weapons Used",
       fill = "Weapon",
       x = NULL,
       y = NULL) +
  scale_fill_manual(values = c("#FFDAB9", "#D2B48C", "#FFA07A", "#FFB6C1", "#FF6347", "#F5DEB3", "#CD853F", "#F4A460", "#E9967A", "#FF4500")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6, family = "Courier"),
        panel.background = element_rect(fill = "#FDF7F6"),
        text = element_text(family = "Courier"),
        legend.text = element_text(family = "Courier"),
        legend.title = element_text(family = "Courier Bold"))
```
In this data, we see that strong arm (hands, feet, or bodily force) were most commonly used of all the weapon types.

## All Weapons Used in the Top 10 Crimes - Did not end up using in presentation!
I wanted to show all tof the weapons used in the top 10 crimes. It ended up being too long to present on so I unfortunately had to cut it out of the slideshow.
```{r, fig.width=17, fig.height=6}
top_10_crimes <- crime %>%
  filter(!is.na(crm_cd_desc)) %>%
  count(crm_cd_desc) %>%
  arrange(desc(n)) %>%
  slice_head(n = 10) %>%
  pull(crm_cd_desc)

color_palette <- c(
  colorRampPalette(brewer.pal(9, "Reds"))(78/3),
  colorRampPalette(brewer.pal(9, "Oranges"))(78/3),
  colorRampPalette(brewer.pal(9, "YlOrBr"))(78/3)
)

crime %>%
  filter(!is.na(weapon_desc) & crm_cd_desc %in% top_10_crimes) %>%
  count(weapon_desc) %>%
  arrange(desc(n)) %>%
  ggplot(aes(x = "", y = n, fill = weapon_desc)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values = color_palette) + 
  coord_polar(theta = "y") +
  labs(title = "Weapons Used in Top 10 Crimes",
       fill = "Weapon",
       x = NULL,
       y = NULL) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6, family = "Courier"),
        panel.background = element_rect(fill = "#FDF7F6"),
        text = element_text(family = "Courier"),
        legend.text = element_text(family = "Courier"),
        legend.title = element_text(family = "Courier Bold"))
```
 
## Timeline of the Crimes

## Most Common Month Crimes Occur - Slide 13
```{r, fig.width=10, fig.height=6}
crime$date_occ <- as.Date(crime$date_occ)

crime %>% 
  mutate(month = lubridate::month(date_occ, label = TRUE)) %>% 
  filter(!is.na(month) & year_occurred >= 2020 & year_occurred <= 2023) %>%
  ggplot(aes(x = month)) +
  geom_bar(fill = "#FFA07A", color = "black") + 
  labs(title = "Crimes Committed by Month (2020-2023)",
       x = "Month",
       y = "Crime Count") +
  theme_bw() +
  theme(text = element_text(family = "Courier"),
        panel.background = element_rect(fill = "#FDF7F6")) +
  facet_wrap(~ year_occurred, ncol = 2)
```

## Most Common Time of Day Crimes Occur - Did not end up using in presentation
```{r, fig.width=10, fig.height=6}
crime$time_occ <- as.numeric(crime$time_occ)
top_10_weapons <- head(table(crime$weapon_desc), 10)
ggplot(crime, aes(x = time_occ, fill = weapon_desc)) +
  stat_count(aes(y = ..count..), geom = "point", size = 3.5, shape = 21, 
             data = subset(crime, weapon_desc %in% names(top_10_weapons))) +
  labs(title = "Crime Occurrence by Time of Day",
       x = "Time of Occurrence",
       y = "Number of Crimes",
       fill = "Weapon Description") +
  scale_fill_manual(values = c("#FFDAB9", "#D2B48C", "#FFA07A", "#FFB6C1", "#FF6347", "#CD853F", "#F4A460", "#E9967A", "#FF4500", "#FFD700", "#E9967A", "#8B4513", "#FFB90F")) +  
  theme_bw() +
  theme(text = element_text(family = "Courier"),
        panel.background = element_rect(fill = "#FDF7F6"),
        legend.text = element_text(family = "Courier"),
        legend.title = element_text(family = "Courier Bold"))
```


# Location - Slide 14

## Where were most of the crimes committed?
Attempted to use ggmaps however there is so much data that it would not load in a clean manner on R Studio. Below is the work for it. I scratched this from the presentation itself and simply showed a map of Los Angeles with a corresponding table.
```{r}
register_stadiamaps("4bb3569f-aecd-42d8-9721-de9f3992e630", write = FALSE)
```

```{r}
crime %>% 
  select(lat, lon) %>% 
  filter(lat != 0 & lon != 0) %>% 
  summary()
```
0.00 was used as a placeholder for locations that were unknown. I removed those to get the accurate boundaries for the maps.
```{r, fig.width=15, fig.height=8}
lat <- c(33.71, 34.33)
long <- c(-118.7, -118.2)
bbox <- make_bbox(long, lat, f = 0.05)
map1 <- get_stadiamap(bbox, maptype = "stamen_terrain", zoom=5)
ggmap(map1) + 
  geom_point(data = crime, aes(lon, lat), size=1, alpha=0.5) +
  labs(x= "Longitude", y= "Latitude", title="Crime Locations")
ggmap(map1)

```

```{r, fig.width= 15, fig.height= 10}
top_10_crime_types <- crime %>%
  group_by(crm_cd_desc) %>%
  summarise(crime_count = n()) %>%
  arrange(desc(crime_count)) %>%
  slice_head(n = 15) %>%
  pull(crm_cd_desc)
top_10_crimes <- crime %>%
  filter(crm_cd_desc %in% top_10_crime_types)
ggmap(map1) + 
  geom_point(data = top_10_crimes, aes(lon, lat), size = 0.05, alpha = 0.5, dpi = 300) +
  labs(x = "Longitude", y = "Latitude", title = "Locations of Top 10 Crimes") +
  theme_bw()
```

## Plotting the Area Names Where the Most Crimes Occured
```{r}
crime %>%
  group_by(area_name) %>%
  summarize(crime_count = n()) %>%
  arrange(desc(crime_count))
```
```{r}
crime_counts <- crime %>%
  group_by(area_name) %>%
  summarise(crime_count = n())

ggplot(crime_counts, aes(x = reorder(area_name, -crime_count), y = crime_count)) +
  geom_point(size = 3, color = "#FFA07A") +
  labs(title = "Crime Counts by Area",
       x = "Area",
       y = "Crime Count") +
  theme_bw() +
  theme(text = element_text(family = "Courier"),
        panel.background = element_rect(fill = "#FDF7F6"),
        legend.text = element_text(family = "Courier"),
        legend.title = element_text(family = "Courier Bold"),
        axis.text.x = element_text(angle = 45, hjust = 1))
```

# Note
After practicing my portion of the presentation, I realized I had too many slides and had to delete a couple. I did still want to show the work I did since it took a while so I kept those unpresented graphs here. 
